---
title: "Visualization"
author: "Jonas Simon Fleck"
date: "17/05/2022"
output: html_document
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

Here we want to show you some ways to visualize the GRN and to plot some metrics that might be interesting to look at for quality control. First, we load an object that already has a GRN inferred.

```{r, results='hide'}
library(tidyverse)
library(patchwork)
library(Pando)
```

```{r, eval=F, echo=T}
muo_data <- read_rds('muo_data.rds')
muo_data@grn
```
```{r, eval=T, echo=F}
muo_data <- read_rds('~/Dropbox/projects/Pando/data/nepi_test.rds')
muo_data@grn
```

## QC and module metrics

As we've shown you in the "Getting started with Pando" vignette, the module discovery step involves applying some filtering criteria on the goodness-of-fit of the fitted models. We can plot this with `plot_gof()`

```{r, eval=T, echo=T}
plot_gof(muo_data, point_size=3)
```

This shows us the explained variance ($R^2$) of the model, as well as the number of variables in each model. We can observe that with more variables, the models will have a better fit. The dashed lines show the applied thresholds for in `find_modules()`.

We can also plot the size of the modules with respect to the number of target genes as well as the number of peaks and regulating TFs per target gene with `plot_module_metrics()`:

```{r, eval=T, echo=T}
plot_module_metrics(muo_data)
```

## Graph visualization

To plot the full GRN as a graph, we found it quite useful to use a UMAP embedding of the nodes. This can be computed with `get_network_graph()`. With the `graph_name` argument, we can give a name to the graph under which we can access it later.

```{r, eval=T, echo=T}
muo_data <- get_network_graph(muo_data, graph_name='umap_graph')
```

To plot it, we use `plot_network_graph()`:

```{r, eval=T, echo=T}
plot_network_graph(muo_data, graph='umap_graph')
```

Here, the nodes are colored and sized based on their centrality in the graph and the edges are colored by the direction of the regulation (inhibitory: grey; activating: orange)

Alternatively, we could also not run UMAP, but simply turn all modules into a graph and visualize it with different graph layouts. For this, we first compute another graph, this time without UMAP:

```{r, eval=T, echo=T}
muo_data <- get_network_graph(
    muo_data, 
    graph_name = 'full_graph', 
    umap_method = 'none'
)
```
\
Again, we can plot it using `plot_network_graph()`, but this time we are specifying different graph layouts. All layouts that can be used through [ggraph](https://ggraph.data-imaginist.com/index.html), can also be used here.

```{r, eval=T, echo=T}
p1 <- plot_network_graph(muo_data, graph='full_graph', layout='fr', text_size=5)
p2 <- plot_network_graph(muo_data, graph='full_graph', layout='drl', text_size=5)
p3 <- plot_network_graph(muo_data, graph='full_graph', layout='eigen', text_size=5)
p4 <- plot_network_graph(muo_data, graph='full_graph', layout='matrix', text_size=5)
(p1 | p2) / (p3 | p4)
```
\
For larger GRNs, we have found that these methods tend to be not overly informative, but for smaller graphs, they can certainly work well. To plot a smaller subgraph of the GRN, we can select a set of genes with the `features` argument of `get_network_graph`:

```{r, eval=T, echo=T}
muo_data <- get_network_graph(
    muo_data, 
    graph_name = 'sub_graph', 
    umap_method = 'none',
    features = c('PAX6', 'RAX', 'GLI2', 'MSX1', 'BAMBI', 'ID4', 'PTCH1', 
                 'OTX2', 'SFRP2', 'SFRP2', 'ID3', 'DLX5', 'EMX2')
)
plot_network_graph(muo_data, graph='sub_graph', layout='fr', color_nodes=F, node_size=5)
```

Sometimes it's also quite useful to plot a subgraph of the GRN centered around one TF, to reveal a hierarchy of regulation. This can be done with the `get_tf_network()` and `plot_tf_network()` functions. The

```{r, eval=T, echo=T}
muo_data <- get_tf_network(muo_data, tf='PAX6', graph='full_graph')
plot_tf_network(muo_data, tf='PAX6')
```

Instead of this circular layout one can also use a more classical tree layout:

```{r, eval=T, echo=T}
plot_tf_network(muo_data, tf='PAX6', circular=F)
```

In the two graphs above, the edges have been pruned to retain only the stronges connection to each terminal node. One can also turn off this behavior with the `keep_all_edges` argument to plot the full graph. Also, there are more arguments that allow changing some aestetics of the plot:

```{r, eval=T, echo=T}
muo_data <- get_tf_network(muo_data, tf='OTX2', graph='full_graph', keep_all_edges = T)
plot_tf_network(
    muo_data, 
    tf = 'OTX2',
    edge_color = c('hotpink', 'turquoise'),
    circular = F,
    edge_width = 1
)
```

## Custom plotting

Finally, if you are familiar with `ggplot` and [`ggraph`](https://ggraph.data-imaginist.com/index.html), then you can of course do some more fancy custom plots. For this, you can access the created graph objects with `NetworkGraph()`.

```{r, eval=T, echo=T}
subgraph <- NetworkGraph(muo_data, graph='sub_graph')
subgraph
```
\
This `tbl_graph` object works with most [igraph](https://igraph.org/) and [`tidygraph`](https://tidygraph.data-imaginist.com/index.html) functions for easy graph manipulation and you can use it to make more customized plots with `ggraph`:

```{r, eval=T, echo=T}
library(ggraph)
ggraph(subgraph, layout='drl') +
    geom_edge_hive(aes(color=estimate), size=2, fill='grey') +
    geom_node_point(aes(size=centrality), shape=21, fill='grey') +
    geom_node_text(aes(label=name), size=4, shape=21, repel=T) +
    theme_void()
```
